Highly Available Transactions: Virtues and Limitations\\(Extended Version)

### 2013-07-04 writeup/evaluation.tex
Start of reviewer comments.

\vspace{.5em}\noindent\textit{Summary.} Many---but not all---TPC-C
transactions are well-served by HATs. The two problematic
transactions, New-Order and Payment, rely on non-monotonic state
update. The former can be modified to ensure ID uniqueness but not
sequential ID ordering, while the latter is inherently
non-monotonic, requiring external compensation or stronger
consistency protocols. Based on these experiences and discussions with
practitioners, we expect that, especially for read-dominated workloads
found in many online services, HAT guarantees will provide useful
semantics for many applications.

@@

\vspace{.5em}\noindent\textit{Implementation.} Our prototype database
is a partially replicated (hash-based, partitioned) key-value backed
by LevelDB and implemented in Java using Apache Thrift. It currently
supports eventual consistency (hereafter, \texttt{eventual};
last-writer-wins RU with standard all-to-all anti-entropy between
replicas) and HAT RC and TA with RC guarantees (hereafter, \texttt{TA}). It also supports
non-HAT operation whereby all operations for a given key are routed to
a (randomly) designated \texttt{master} replica for each key
(hereafter, \texttt{master}; i.e., PNUTS~\cite{pnuts}) as well as
distributed two-phase locking. Servers are durable: they synchronously
write to LevelDB before responding to client requests, while new
writes in TA are sent to a write-ahead log, which is synchronously
flushed to disk.

>>>>

\vspace{.5em}\noindent\textit{Summary.} Many---but not all---TPC-C
transactions are well-served by HATs. The two problematic
transactions, New-Order and Payment, rely on non-monotonic state
update. The former can be modified to ensure ID uniqueness but not
sequential ID ordering, while the latter is inherently non-monotonic,
requiring external compensation or stronger consistency
protocols. Based on these experiences and discussions with
practitioners, we believe that HAT guarantees can provide useful
semantics for a large class of application functionality, while a
(possibly small) subset of operations will require stronger,
unavailable properties.

@@

\vspace{.5em}\noindent\textit{Implementation.} Our prototype database
is a partially replicated (hash-based, partitioned) key-value backed
by LevelDB and implemented in Java using Apache Thrift. It currently
supports eventual consistency (hereafter, \texttt{eventual};
last-writer-wins RU with standard all-to-all anti-entropy between
replicas) and HAT RC and TA with RC guarantees (hereafter,
\texttt{TA}). It also supports non-HAT operation whereby all
operations for a given key are routed to a (randomly) designated
\texttt{master} replica for each key (guaranteeing single-key
linearizability, as in Gilbert and Lynch's CAP Theorem
proof~\cite{gilbert-cap} and in PNUTS~\cite{pnuts}'s
\texttt{read-latest} operation; hereafter, \texttt{master}) as well as
distributed two-phase locking. Servers are durable: they synchronously
write to LevelDB before responding to client requests, while new
writes in TA are sent to a write-ahead log, which is synchronously
flushed to disk.

### 2013-07-04 writeup/hat-analysis.tex
Start of reviewer comments.

Perplexingly, discussions of TA are absent from existing treatments of
weak isolation. This is perhaps again due to the single-node context
in which prior work was developed: on a single server (or a fully
replicated database), TA is achievable via lightweight locking and/or
local concurrency control over data items~\cite{gstore}. In contrast,
in a distributed environment, TA over arbitrary groups of
non-colocated items is considerably more difficult to achieve with
high availability: servers need to ensure that, regardless of replica
selection, all of a client's accesses will read a sufficiently
up-to-date/``consistent'' set of data items. However, replicas do not
need to agree on when to choose to reveal sets of values to clients, which would require consensus. Instead, as we see below, TA
only requires the equivalent of reliable broadcast (with some
additional client metadata) and is achievable in HAT systems.

@@

\vspace{.5em}\noindent\textbf{{Writes Follow Reads}} requires that, if
a session observes an effect of transaction $T_1$ and subsequently
commits transaction $T_2$, then another session can only observe
effects of $T_2$ if it can also observe $T_1$'s effects (or later
values that supersede $T_1$'s).  Any order on transactions should
respect the reads-from order.\vspace{.5em}

The above guarantees can be achieved by forcing servers to wait to
reveal new writes until each write's respective dependencies are
visible on all replicas. This mechanism effectively ensures that all
clients read from a globally agreed upon lower bound on the versions
written. This is highly available as a client will never block due to
inability to find a server with a sufficiently up-to-date version of a
data item. However, it does not imply that transactions will read
their own writes or, in the presence of partitions, make forward
progress through the version history. The problem is that under
non-sticky high availability, a system must handle the possibility
that, under a partition, an unfortunate client will be forced to issue
its next requests against a partitioned, out-of-date server.

@@

%\vspace{.5em}\noindent{\textbf{Durability}} A client requiring that its
%transactions' effects survive $F$ server faults requires at least
%$F+1$ non-failing replicas; this affects availability.

@@

Their need to prevent Lost Update means that Consistent Read, Snapshot
Isolation, and Cursor Stability guarantees are all unavailable.
Repeatable Read (defined by Gray~\cite{gray-isolation}, Berenson et
al.~\cite{ansicritique}, and Adya~\cite{adya}) and One-Copy
Serializability need to prevent both Lost Update and Write Skew. These
prevention requirements mean that these guarantees are inherently
unavailable.

@@

>>>>

Perplexingly, discussions of TA are absent from existing treatments of
weak isolation. This is perhaps again due to the single-node context
in which prior work was developed: on a single server (or a fully
replicated database), TA is achievable via lightweight locking and/or
local concurrency control over data items~\cite{gstore,
  kemme-thesis}. In contrast, in a distributed environment, TA over
arbitrary groups of non-colocated items is considerably more difficult
to achieve with high availability: servers need to ensure that,
regardless of replica selection, all of a client's accesses will read
a sufficiently up-to-date/``consistent'' set of data items. However,
replicas do not need to agree on when to choose to reveal sets of
values to clients, which would require consensus. Instead, as we see
below, TA only requires the equivalent of reliable broadcast (with
some additional client metadata) and is achievable in HAT systems.

@@

\vspace{.5em}\noindent\textbf{{Writes Follow Reads}} requires that, if
a session observes an effect of transaction $T_1$ and subsequently
commits transaction $T_2$, then another session can only observe
effects of $T_2$ if it can also observe $T_1$'s effects (or later
values that supersede $T_1$'s); this corresponds to Lamport's
``happens-before'' relation~\cite{lamportclocks}.  Any order on
transactions should respect the transitive reads-from
order.\vspace{.5em}

The above guarantees can be achieved by forcing servers to wait to
reveal new writes (say, by buffering them in separate local storage)
until each write's respective dependencies are visible on all
replicas. This mechanism effectively ensures that all clients read
from a globally agreed upon lower bound on the versions written. This
is highly available as a client will never block due to inability to
find a server with a sufficiently up-to-date version of a data
item. However, it does not imply that transactions will read their own
writes or, in the presence of partitions, make forward progress
through the version history. The problem is that under non-sticky high
availability, a system must handle the possibility that, under a
partition, an unfortunate client will be forced to issue its next
requests against a partitioned, out-of-date server.

@@

@@

Their need to prevent Lost Update means that Consistent Read, Snapshot
Isolation (including Parallel Snapshot Isolation~\cite{walter}), and
Cursor Stability guarantees are all unavailable.  Repeatable Read
(defined by Gray~\cite{gray-isolation}, Berenson et
al.~\cite{ansicritique}, and Adya~\cite{adya}) and One-Copy
Serializability~\cite{1sr} need to prevent both Lost Update and Write Skew. These
prevention requirements mean that these guarantees are inherently
unavailable.

@@

\subsubsection{Durability}

A client requiring that its transactions' effects survive $F$ server
faults requires at least $F+1$ non-failing replicas; this affects
availability and, according to the Gilbert and Lynch definition we
have adopted, $F>1$ fault tolerance is not achievable with high
availability.

### 2013-07-04 writeup/hat-appendix.tex
Start of reviewer comments.

\begin{algorithmic}
\caption*{\textbf{Algorithm} HAT Transactional Atomicity}\vspace{.5em}
\State{\textit{\textbf{Shared Data Types}}\\\vspace{.5em}
\texttt{timestamp}: unique, per transaction identifier\\
$\texttt{write}:[\texttt{key:}k, \texttt{value}:v, \texttt{timestamp}:ts, \texttt{set<key>}:tx\_keys]$\\\hrulefill}\vspace{.5em}

@@

\Procedure{put}{\texttt{write}:$w$}
  \State \texttt{pending}.add($w$)
  \For{key $k_t \in w.tx\_keys$}
  \State \{all replicas for $sib$\}.notify($w.ts$)
  \EndFor

@@

  \Procedure{notify}{\texttt{timestamp}:$ts$}
  \State \texttt{acks}.get($ts$).increment()
  \If{all \texttt{acks} received for all replicas for $ts$'s $tx\_keys$}
  \State \texttt{good}.add($w$)
  \State \texttt{pending}.remove($w$)  
  \EndIf  
  \EndProcedure\vspace{.5em}

@@

  \State $w\_{ret}$ = (available replica for $k$).get($k$, \texttt{required}.get($k$))
  \For {$(tx_{key} \in w_{ret}.tx\_keys$}
  \If {$w_{ret}.ts > \texttt{required}.get(tx_{key})$}
  \State $\texttt{required}$.put$(tx_{key}, w_{ret}.ts)$
  \EndIf
  \EndFor

>>>>

\begin{algorithmic}
\caption*{\textbf{Algorithm} HAT Transactional Atomicity}\vspace{.5em}
\State{\textit{\textbf{Shared Data Types}}\\\vspace{.5em}
\texttt{timestamp}: unique, per transaction identifier\\
$\texttt{write}:[\texttt{key:}k, \texttt{value}:v, \texttt{timestamp}:ts, \texttt{set<key>}:sibs]$\\\hrulefill}\vspace{.5em}

@@

\Procedure{put}{\texttt{write}:$w$}
  \State \texttt{pending}.add($w$)
  \For{key $k_t \in w.sibs$}
  \State \{all replicas for $sib$\}.notify($w.ts$)
  \EndFor

@@

  \Procedure{notify}{\texttt{timestamp}:$ts$}
  \State \texttt{acks}.get($ts$).increment()
  \If{all \texttt{acks} received for all replicas for $ts$'s $sibs$}
  \State \texttt{good}.add($w$)
  \State \texttt{pending}.remove($w$)  
  \EndIf  
  \EndProcedure\vspace{.5em}

@@

  \State $w\_{ret}$ = (available replica for $k$).get($k$, \texttt{required}.get($k$))
  \For {$(tx_{key} \in w_{ret}.sibs$}
  \If {$w_{ret}.ts > \texttt{required}.get(tx_{key})$}
  \State $\texttt{required}$.put$(tx_{key}, w_{ret}.ts)$
  \EndIf
  \EndFor

### 2013-07-04 writeup/model.tex
Start of reviewer comments.

Informally, highly available algorithms ensure ``always on'' operation
and guaranteed low latency. If users of a highly available system are
able to contact a (set of) server(s) in a system, they are guaranteed
a response; this means servers will not need to communicate with any
others. If servers are partitioned from one another, they do not need
to stall in order to provide clients a ``safe'' response to
operations. This lack of fast-path coordination also means that a
highly available system also provides low latency~\cite{abadi-pacelc};
in a wide-area setting, clients of a highly available system need not
wait for cross-datacenter communication. To properly describe whether
a \textit{transactional} system is highly available, we need to
describe what servers a client must contact as well as what kinds of
responses a server can provide, especially given the possibility of
aborts.

@@

Traditionally, a system provides {\em high availability} if every user
that can contact a correct (non-failing) server eventually receives a
response from that server, even in the presence of arbitrary,
indefinitely long network partitions between
servers~\cite{gilbert-cap}.\footnote{Under this definition, systems
  that require a majority of servers to be online is not
  available. Similarly, a system which ensures that clients receive a
  response with high (but not perfect) probability is not available.}
As in a standard distributed database, designated servers might
perform operations for different data items; a server that can handle
an operation for a given data item is called a \textit{replica} for
that item.\footnote{There is a further distinction between a
  \textit{fully replicated} system, in which all servers are replicas
  for all data items, and a \textit{partially replicated} system, in
  which at least one server acts as a replica for only a subset of all
  data items. For generality, and, for applicability to many modern
  ``sharded'' or ``partitioned'' data storage systems~\cite{
   bigtable, pnuts,  spanner, dynamo, hstore}, we consider partially
  replicated systems.}

In addition to high availability, which allows operations on any replica,
distributed algorithms often assume a model in which clients always
contact the same (sets of) replica(s) across subsequent operations. As
we will discuss in Section~\ref{sec:hats}, clients can ensure
continuity between operations (e.g., reading their prior updates to a
data item) by maintaining affinity or ``stickiness'' with a server or
set of servers~\cite{vogels-defs}. Simultaneously maintaining
availability and stickiness requires fate sharing between the client
and its server: if a client's sticky servers fail or if the client is
partitioned from its sticky servers, then it may face
unavailability. We say that a system provides \textit{sticky
  availability} if, whenever a client's operation is executed against
a correct server that has observed all of its prior operations, it
eventually receives a response, even in the presence of indefinitely
long partitions. A client may choose to become sticky available by
acting as a server itself; for example, a client might cache its reads
and writes~\cite{bolton, sessionguarantees, swift}. Any guarantee
achievable in a highly available system is achievable in a sticky high
availability system but not vice-versa.

@@

Until now, we have considered single-object, single-operation
availability. This is standard in the distributed systems literature
(e.g., distributed register models such as linearizability all concern
single objects~\cite{herlihy-art}), yet the database literature
largely focuses on transactions: groups of multiple operations over
multiple objects. Accordingly, by itself, replica availability is
insufficient to describe availability guarantees for
transactions. Additionally, given the choice of \textit{commit} and
\textit{abort} responses---which signal transaction success or failure
to a client---we must take care in defining transactional
availability.

If a client wishes to execute a transaction that performs operations
on multiple data items, then the client must be able to contact and
receive a response from at least one replica for each data item. This
may result in ``lower availability'' than a non-transactional
requirement. Additionally, given the possibility of system-initiated
aborts, we need to ensure useful forward progress: a system can
trivially guarantee clients a response by always aborting all
transactions. However, this is an unsatisfactory system because
nothing good (transaction commit) ever happens; we should require a
\textit{liveness} property~\cite{transaction-liveness}. We cannot
guarantee that every transaction will commit---transactions may choose
to abort themselves---but we need to make sure that the system will
not indefinitely abort transactions on its own volition. We call a
transaction abort due to a transaction's own choosing (e.g., as an
operation of the transaction itself or due to a would-be declared
integrity constraint violation) an \textit{internal abort} and an
abort due to system implementation or operation an \textit{external abort}.

We say that a system provides \textit{transactional availability} if,
given replica availability for every data item in a transaction, the
transaction eventually commits (possibly after multiple client
retries) or internally aborts~\cite{hat-hotos}. For example, a system
will violate transactional availability if a client that can contact
two servers for each of the data items in its transaction does not
eventually commit in the absence of internal aborts.

>>>>

Informally, highly available algorithms ensure ``always on'' operation
and, as a side effect, guarantee low latency. If users of a highly
available system are able to contact a (set of) server(s) in a system,
they are guaranteed a response; this means servers will not need to
communicate with any others. If servers are partitioned from one
another, they do not need to stall in order to provide clients a
``safe'' response to operations. This lack of fast-path coordination
also means that a highly available system also provides low
latency~\cite{abadi-pacelc}; in a wide-area setting, clients of a
highly available system need not wait for cross-datacenter
communication. To properly describe whether a \textit{transactional}
system is highly available, we need to describe what servers a client
must contact as well as what kinds of responses a server can provide,
especially given the possibility of aborts.

@@

Traditionally, a system provides {\em high availability} if every user
that can contact a correct (non-failing) server eventually receives a
response from that server, even in the presence of arbitrary,
indefinitely long network partitions between
servers~\cite{gilbert-cap}.\footnote{Under this definition, systems
  that require a majority of servers to be online is not
  available. Similarly, a system which ensures that clients receive a
  response with high (but not perfect) probability is not
  available. This is a stringent definition but matches the
  assumptions made by the CAP Theorem~\cite{gilbert-cap} and
  guarantees low latency~\cite{abadi-pacelc}.} As in a standard
distributed database, designated servers might perform operations for
different data items; a server that can handle an operation for a
given data item is called a \textit{replica} for that
item.\footnote{There is a further distinction between a \textit{fully
    replicated} system, in which all servers are replicas for all data
  items, and a \textit{partially replicated} system, in which at least
  one server acts as a replica for only a subset of all data
  items. For generality, and, for applicability to many modern
  ``sharded'' or ``partitioned'' data storage systems~\cite{ bigtable,
    pnuts, spanner, dynamo, hstore}, we consider partially replicated
  systems.}

\subsection{Sticky Availability}

In addition to high availability, which allows operations on any
replica, distributed algorithms often assume a model in which clients
always contact the same logical (sets of) replica(s) across subsequent
operations. As we will discuss in Section~\ref{sec:hats}, clients can
ensure continuity between operations (e.g., reading their prior
updates to a data item) by maintaining affinity or ``stickiness'' with
a server or set of servers~\cite{vogels-defs}. Simultaneously
maintaining availability and stickiness requires fate sharing between
the client and its server: if a client's sticky servers fail or if the
client is partitioned from its sticky servers, then it may face
unavailability. We say that a system provides \textit{sticky
  availability} if, whenever a client's operation on an item is
executed against a correct replica that has observed all of its prior
operations for that item, it eventually receives a response, even in
the presence of indefinitely long partitions. A client may choose to
become sticky available by acting as a server itself; for example, a
client might cache its reads and writes~\cite{bolton,
  sessionguarantees, swift}. Any guarantee achievable in a highly
available system is achievable in a sticky high availability system
but not vice-versa.

@@

Until now, we have considered single-object, single-operation
availability. This is standard in the distributed systems literature
(e.g., distributed register models such as linearizability all concern
single objects~\cite{herlihy-art}), yet the database literature
largely focuses on transactions: groups of multiple operations over
multiple objects. Accordingly, by itself, traditional definitions of
high availability are insufficient to describe availability guarantees
for transactions. Additionally, given the choice of \textit{commit}
and \textit{abort} responses---which signal transaction success or
failure to a client---we must take care in defining transactional
availability.

If a client wishes to execute a transaction that performs operations
on multiple data items, then the client must be able to contact and
receive a response from at least one replica for each data item. We
say that a transaction has \emph{replica availability} if it can
contact at least one replica for every item it attempts to
access. This may result in ``lower availability'' than a
non-transactional requirement. Additionally, given the possibility of
system-initiated aborts, we need to ensure useful forward progress: a
system can trivially guarantee clients a response by always aborting
all transactions. However, this is an unsatisfactory system because
nothing good (transaction commit) ever happens; we should require a
\textit{liveness} property~\cite{transaction-liveness}. We cannot
guarantee that every transaction will commit---transactions may choose
to abort themselves---but we need to make sure that the system will
not indefinitely abort transactions on its own volition. We call a
transaction abort due to a transaction's own choosing (e.g., as an
operation of the transaction itself or due to a would-be declared
integrity constraint violation) an \textit{internal abort} and an
abort due to system implementation or operation an \textit{external
  abort}. We say that a system provides \textit{transactional
  availability} if, given replica availability for every data item in
a transaction, the transaction eventually commits (possibly after
multiple client retries) or internally aborts~\cite{hat-hotos}.

### 2013-07-04 writeup/motivation.tex
fix labels
Start of reviewer comments.

Several recent studies rigorously quantify partition behavior. A 2011
study of several Microsoft datacenters found a mean of 40.8 network
link failures per day (95th percentile: 136), with a median time to
repair of around five minutes (and up to one week). Perhaps
surprisingly, provisioning redundant networks only reduces impact of
failures by up to 40\%, meaning network providers cannot easily
curtail partition behavior~\cite{sigcomm-dc}. A 2010 study of over 200
wide-area routers found an average of 16.2--302.0 failures per link
per year with an average annual downtime of 24--497 minutes per link
per year (95th percentile at least 34 hours)~\cite{sigcomm-wan}. In
HP's managed enterprise networks, WAN, LAN, and connectivity problems
account for 28.1\% of all customer support tickets while 39\% of
tickets relate to network hardware.  The median incident duration for
highest priority tickets ranges from 114--188 minutes and up to a full
day for all tickets~\cite{turner2012failure}. Other studies confirm
these results, showing median time between failures over a WAN network
of approximately 3000 seconds with a median time to repair between 2
and 1000 seconds~\cite{ip-backbone-failures} as well as frequent path
routing failures on the Internet~\cite{labovitz-failures}. Isolating,
quantifying, and accounting for these network failures is an area of
active research in networking community~\cite{uw-failure-networks}.

@@

\subfloat[Across \texttt{us-east} AZs]{
  \makebox[.2\textwidth]{
    \begin{tabular}{|c|c|c|}\hline
 & \multicolumn{1}{c}{C} & \multicolumn{1}{c|}{D}\\\hline
B & \colorbox{min-lat-color}{1.08} & 3.12 \\
C & & \colorbox{max-lat-color}{3.57}  \\
\hline
  \end{tabular}}
 }
\subfloat[Within \texttt{us-east-b} AZ] {
  \makebox[.25\textwidth]{
  \begin{tabular}{|c|c|c|}\hline
 & \multicolumn{1}{c}{H2} & \multicolumn{1}{c|}{H3}\\\hline
H1  & 0.55   & \colorbox{max-lat-color}{0.56} \\
H2 &  & \colorbox{min-lat-color}{0.50}  \\
\hline
  \end{tabular}
 }}

@@

In real deployments, messages travel slower than the speed of light
due to routing, congestion, and server-side overheads. To illustrate
the difference between intra-datacenter, inter-datacenter, and
inter-planetary networks, we performed a measurement study of network
behavior on Amazon's EC2, a widely used public compute cloud. We
measured one week of ping times (i.e., round-trip times, or
RTTs) between all seven EC2 geographic ``regions,'' across three
``availability zones'' (closely co-located datacenters), and within a
single ``availability zone'' (datacenter), at a granularity of 1s
(dataset to be released). We summarize the results of our network
measurement study in Table~\ref{table:rtt}. On average,
intra-datacenter communication (Table 1a) is between 1.82 and 6.38
times faster than across geographically co-located datacenters (Table
1b) and between 40 and 647 times faster than across geographically
distributed datacenters (Table 1c). The cost of wide-area
communication exceeds the speed of light: for example, while a
speed-of-light RTT from S\~{a}o Paulo to Singapore RTT is 106.7ms,
ping packets incur an average 362.8ms RTT (95th percentile: 649ms). As
shown in Figure~\ref{fig:rtt}, the distribution of latencies varies
between links, but the trend is clear: remote communication has a
substantial cost.

>>>>

Several recent studies rigorously quantify partition behavior. A 2011
study of several Microsoft datacenters observed over 13,300 network
failures with end-user impact, with an estimated median 59,000 packets
lost per failure. The study found a mean of 40.8 network link failures
per day (95th percentile: 136), with a median time to repair of around
five minutes (and up to one week). Perhaps surprisingly, provisioning
redundant networks only reduces impact of failures by up to 40\%,
meaning network providers cannot easily curtail partition
behavior~\cite{sigcomm-dc}. A 2010 study of over 200 wide-area routers
found an average of 16.2--302.0 failures per link per year with an
average annual downtime of 24--497 minutes per link per year (95th
percentile at least 34 hours)~\cite{sigcomm-wan}. In HP's managed
enterprise networks, WAN, LAN, and connectivity problems account for
28.1\% of all customer support tickets while 39\% of tickets relate to
network hardware.  The median incident duration for highest priority
tickets ranges from 114--188 minutes and up to a full day for all
tickets~\cite{turner2012failure}. Other studies confirm these results,
showing median time between connectivity failures over a WAN network
of approximately 3000 seconds with a median time to repair between 2
and 1000 seconds~\cite{ip-backbone-failures} as well as frequent path
routing failures on the Internet~\cite{labovitz-failures}. Finally,
Kingsbury and Bailis have recently collected an informal survey of
network failures~\cite{aphyr-post}. Isolating, quantifying, and
accounting for these network failures is an area of active research in
networking community~\cite{uw-failure-networks}.

@@

\subfloat[Within \texttt{us-east-b} AZ] {
  \makebox[.2\textwidth]{
 \begin{tabular}{|c|c|c|}\hline
 & \multicolumn{1}{c}{H2} & \multicolumn{1}{c|}{H3}\\\hline
H1  & 0.55   & \colorbox{max-lat-color}{0.56} \\
H2 &  & \colorbox{min-lat-color}{0.50}  \\
\hline
  \end{tabular}}
 }
\subfloat[Across \texttt{us-east} AZs]{
  \makebox[.25\textwidth]{
    \begin{tabular}{|c|c|c|}\hline
 & \multicolumn{1}{c}{C} & \multicolumn{1}{c|}{D}\\\hline
B & \colorbox{min-lat-color}{1.08} & 3.12 \\
C & & \colorbox{max-lat-color}{3.57}  \\
\hline
  \end{tabular}}
}

@@

In real deployments, messages travel slower than the speed of light
due to routing, congestion, and server-side overheads. To illustrate
the difference between intra-datacenter, inter-datacenter, and
inter-planetary networks, we performed a measurement study of network
behavior on Amazon's EC2, a widely used public compute cloud. We
measured one week of ping times (i.e., round-trip times, or RTTs)
between all seven EC2 geographic ``regions,'' across three
``availability zones'' (closely co-located datacenters), and within a
single ``availability zone'' (datacenter), at a granularity of 1s
(dataset to be released). We summarize the results of our network
measurement study in Table~\ref{table:rtt}. On average,
intra-datacenter communication (Table 1a) is between 1.82 and 6.38
times faster than across geographically co-located datacenters (Table
1b) and between 40 and 647 times faster than across geographically
distributed datacenters (Table 1c). The cost of wide-area
communication exceeds the speed of light: for example, while a
speed-of-light RTT from S\~{a}o Paulo to Singapore RTT is 106.7ms,
ping packets incur an average 362.8ms RTT (95th percentile: 649ms). As
shown in Figure~\ref{fig:rtt}, the distribution of latencies varies
between links, but the trend is clear: remote communication has a
substantial cost. Like failures, quantifying and delays communication
delays is an active area of research in the networking
community~\cite{bobtail}.
